---
title: "Data_Processing_MSstatsTMT"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data_Processing_MSstatsTMT}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=TRUE}
devtools::load_all(".")
```

Library setup

```{r setup}
library(MSstatsThermalProfiler)
library(MSstats)
library(MSstatsConvert)
library(MSstatsTMT)
library(tidyr)
library(stringr)
```

# Data Import

### Import option 1: Read in your own file

**Use getwd(your_working_directory) to get a folder where your PSM data is stored**

```{r}
#WD<-setwd("~/CS7290/MSstatsThermalProfiler/data")
#PSMsample_HumanData<-readPSMs(WD)
```

### Import option 2: Use the example file

**load the peptide spectrum match (PSM) example file into R**

```{r}
data("PSMsample_HumanData", package = "MSstatsThermalProfiler")

```

### Verify that the PSM data is loaded

```{r}
head(PSMsample_HumanData)
```

## Remove carrier channel from the data

```{r}
PSMsample_HumanData<-PSMsample_HumanData[,!stringr::str_detect(names(PSMsample_HumanData),"131C")]
```

# Annotation File Generation

### Rename the columns from the PSM file and generate annotation file

**Generate annotation file**

```{r}
pd_annotation<-Annotation2MSstatsTMT(PSMsample_HumanData,solvent="DMSO",temps=Channel2Temps_HumanData,reference="126")

```

# Data processing with MSstatsTMT

1.  Import the data into MSStatsTMT

```{r}

processed.input <- MSstatsTMT::PDtoMSstatsTMTFormat(PSMsample_HumanData,
                                                    pd_annotation,
                                                    which.proteinid = "Master Protein Accessions",
                                                    useNumProteinsColumn = TRUE,
                                                    useUniquePeptide = FALSE,
                                                    rmPSM_withfewMea_withinRun = TRUE,
                                                    rmProtein_with1Feature = TRUE,
                                                    summaryforMultipleRows = max,
                                                    use_log_file = FALSE,
                                                    append=FALSE,
                                                    verbose=TRUE)
```

2.  Summarize the peptide data into proteins

```{r}
summarised.proteins <- proteinSummarization(data = processed.input,
                                method = "msstats",
                                global_norm = FALSE,
                                reference_norm = TRUE,
                                MBimpute=FALSE,
                                use_log_file=FALSE,
                                remove_norm_channel=TRUE)
```

3.  Create a contrast matrix to measure the average treatment effect

    ```{r}

    # perform contrast comparison
    groups <- unique(summarised.proteins$FeatureLevelData$Condition)

    #View the groups
    groups

    #Select the groups without Norm channel or carrier channel
    groups<-groups[!stringr::str_detect(groups,"Norm")]

    groups<-droplevels(groups,"Norm")


    # 'Norm' will be removed during tesing and should be not considered in the contrast
    comparison <-matrix(c(rep(-1/9,9),rep(1/9,9)),nrow=1)


    comparison_names <- c("127N_vehicle",
                          "127C_vehicle",
                          "128N_vehicle",
                          "128C_vehicle",
                          "129N_vehicle",
                          "129C_vehicle",
                          "130N_vehicle",
                          "130C_vehicle",
                          "131N_vehicle",
                          
                          "127N_treated",
                          "127C_treated",
                          "128N_treated",
                          "128C_treated",
                          "129N_treated",
                          "129C_treated",
                          "130N_treated",
                          "130C_treated",
                          "131N_treated")

    # Set the column names
    colnames(comparison)<- comparison_names
    # Set the names of each row
    row.names(comparison)<- "treated-vehicle"

    ```

4.  Fit linear mixed models with groupComparisons from MSstatsTMT

```{r}
ATE<-groupComparisonTMT(
  summarised.proteins,
  contrast.matrix = comparison,
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL
)
```
