---
title: "Simulations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MSstatsThermalProfiler)
library(MSstatsTMT)
```

# MS-Stats Group Comparison

First we need to convert the simulated data sets to the format used by MSstatsTMT

```{r,eval=FALSE}

sigmoid_truepositive_simdata_msstat = list(ProteinLevelData=Sim2MSstatsTMT(sigmoid_truepositive_simdata$ProteinLevelData))

sigmoid_truenegative_simdata_msstat = list(ProteinLevelData=Sim2MSstatsTMT(sigmoid_truenegative_simdata$ProteinLevelData))

#convert condition to factor

sigmoid_truepositive_simdata_msstat$ProteinLevelData$Condition = as.factor(sigmoid_truepositive_simdata_msstat$ProteinLevelData$Condition)

sigmoid_truenegative_simdata_msstat$ProteinLevelData$Condition = as.factor(sigmoid_truenegative_simdata_msstat$ProteinLevelData$Condition)

```

Next we make the contrast matrix to obtain the ATE (averaged over the channels/temps) in MS-Stats

```{r,eval=FALSE}

sigmoid_truepos_compmat = make_contrast_matrix(sigmoid_truepositive_simdata_msstat)

sigmoid_trueneg_compmat = make_contrast_matrix(sigmoid_truenegative_simdata_msstat)

```

Now we are ready for groupComparisonsTMT(). We turn off logging and do not save the fitted model here. Also no moderation is used for the first go. 

## True Positives Result MS-stats

```{r,eval=FALSE}

start=proc.time()
sigmoid_truepos_MSstat_result = groupComparisonTMT(data=sigmoid_truepositive_simdata_msstat,contrast.matrix=sigmoid_truepos_compmat,
                                                   moderated=FALSE,save_fitted_models = FALSE,use_log_file = FALSE)
end=proc.time()
print(end-start)

```


```{r}

start=proc.time()
sigmoid_trueneg_MSstat_result = groupComparisonTMT(data=sigmoid_truenegative_simdata_msstat,contrast.matrix=sigmoid_trueneg_compmat,
                                                   moderated=FALSE,save_fitted_models = FALSE,use_log_file = FALSE)
end=proc.time()
print(end-start)

```

```{r}
start=proc.time()
sigmoid_truepos_MSstat_result_mod = groupComparisonTMT(data=sigmoid_truepositive_simdata_msstat,contrast.matrix=sigmoid_truepos_compmat,
                                                   moderated=TRUE,save_fitted_models = FALSE,use_log_file = FALSE)
end=proc.time()
print(end-start)

```

```{r}

start=proc.time()
sigmoid_trueneg_MSstat_result_mod = groupComparisonTMT(data=sigmoid_truenegative_simdata_msstat,contrast.matrix=sigmoid_trueneg_compmat,
                                                   moderated=TRUE,save_fitted_models = FALSE,use_log_file = FALSE)
end=proc.time()
print(end-start)

```



## Investigate False Negative and ATE and delta Tm relation 

```{r}
deltaTm_truepos = find_deltaTm(sigmoid_truepositive_simdata)
```


```{r,eval=FALSE}

sigmoid_truepos_MSstat_result$ComparisonResult = sigmoid_truepos_MSstat_result$ComparisonResult |> dplyr::inner_join(sigmoid_truepositive_simdata$params,by="Protein") |> dplyr::mutate(ATE_bin = cut(ATE,breaks=c(seq(0,0.06,0.01),0.08,0.1,0.16)))

```

```{r}

sigmoid_truepos_MSstat_result$ComparisonResult = sigmoid_truepos_MSstat_result$ComparisonResult |> dplyr::inner_join(deltaTm_truepos |> dplyr::select(Protein,delta_Tm),by="Protein")


```


```{r}
sigmoid_truepos_MSstat_result$ComparisonResult |> group_by(ATE_bin) |> summarise(N=n(),Sens = mean(pvalue<0.05))
```

```{r}
sigmoid_truepos_MSstat_result$ComparisonResult |> dplyr::mutate(delta_Tm_bin=cut(delta_Tm,breaks=c(-4,-1,0,1,2,3,5))) |> group_by(delta_Tm_bin) |> summarise(N=n(),Sens = mean(pvalue<0.05))
```


## Try with only treated vs vehicle, no channels

```{r}
sigmoid_truepositive_simdata_msstat_nochannel = sigmoid_truepositive_simdata_msstat

sigmoid_truepositive_simdata_msstat_nochannel$ProteinLevelData = sigmoid_truepositive_simdata_msstat_nochannel$ProteinLevelData |>
  mutate(Group=ifelse(stringr::str_detect(Group,"treated"),"treated","vehicle"),
         Condition = Group)

sigmoid_truenegative_simdata_msstat_nochannel = sigmoid_truenegative_simdata_msstat
sigmoid_truenegative_simdata_msstat_nochannel$ProteinLevelData = sigmoid_truenegative_simdata_msstat_nochannel$ProteinLevelData |> mutate(Group=ifelse(stringr::str_detect(Group,"treated"),"treated","vehicle"),
         Condition = Group)

```


```{r}

start=proc.time()
sigmoid_truepos_MSstat_result_nochannel = groupComparisonTMT(data=sigmoid_truepositive_simdata_msstat_nochannel,
                                                   moderated=FALSE,save_fitted_models = FALSE,use_log_file = FALSE)
end=proc.time()
print(end-start)
```

```{r}
start=proc.time()
sigmoid_trueneg_MSstat_result_nochannel = groupComparisonTMT(data=sigmoid_truenegative_simdata_msstat_nochannel,
                                                   moderated=FALSE,save_fitted_models = FALSE,use_log_file = FALSE)
end=proc.time()
print(end-start)
```

# TPP Comparisons

## Running TPP-NPARC implementation

```{r}

TPP_truepos_results = TPP_NPARC_calc(sigmoid_truepositive_simdata,method="NPARC",DF=5,df.temps=Channel2Temps_HumanData,CARRIER=TRUE)

```


```{r}
TPP_trueneg_results = TPP_NPARC_calc(sigmoid_truenegative_simdata,method="NPARC",DF=5,df.temps=Channel2Temps_HumanData,CARRIER=TRUE)
```


```{r}
compute_PPV = function(truepos_pvals,trueneg_pvals,prevalance=0.10){
  sens = mean(truepos_pvals <= 0.05) #sensitivity is true positive rate 
  spec = mean(trueneg_pvals > 0.05 ) #specificity is true negative rate
  
  PPV = (sens*prevalance)/((sens*prevalance) + (1-spec)*((1-prevalance)))
  result = dplyr::tibble(sens=sens,spec=spec,PPV=PPV)
  return(result)
}
```


## Partition by ATE and delta Tm

```{r}

deltaTm_truepos = find_deltaTm(sigmoid_truepositive_simdata)

MSstat_truepos_results = sigmoid_truepos_MSstat_result$ComparisonResult |> dplyr::inner_join(sigmoid_truepositive_simdata$params,by="Protein") |> dplyr::mutate(ATE_bin = cut(ATE,breaks=c(seq(0,0.06,0.01),0.08,0.1,0.16)))

MSstat_truepos_results = MSstat_truepos_results |> dplyr::inner_join(deltaTm_truepos |> dplyr::select(Protein,delta_Tm),by="Protein")

```

```{r}
MSstat_truepos_results |> dplyr::group_by(ATE_bin) |> dplyr::summarise(N=dplyr::n(),sens = mean(pvalue<0.05)) |> dplyr::ungroup() |>
  dplyr::mutate(spec = mean(sigmoid_trueneg_MSstat_result$ComparisonResult$pvalue>0.05), prevalance = 0.10,
                PPV = (sens*prevalance)/((sens*prevalance) + (1-spec)*((1-prevalance))))
```


```{r}
MSstat_truepos_results |> dplyr::mutate(delta_Tm_bin=cut(delta_Tm,breaks=c(-4,-1,0,1,2,3,5))) |> dplyr::group_by(delta_Tm_bin) |> dplyr::summarise(N=dplyr::n(),sens = mean(pvalue<0.05)) |> dplyr::ungroup() |>
  dplyr::mutate(spec = mean(sigmoid_trueneg_MSstat_result$ComparisonResult$pvalue>0.05), prevalance = 0.10,
                PPV = (sens*prevalance)/((sens*prevalance) + (1-spec)*((1-prevalance))))
```

```{r}
TPP_truepos_results = TPP_truepos_results |> dplyr::rename(Protein=uniqueID) |> dplyr::inner_join(sigmoid_truepositive_simdata$params,by="Protein") |> dplyr::mutate(ATE_bin = cut(ATE,breaks=c(seq(0,0.06,0.01),0.08,0.1,0.16)))

TPP_truepos_results = TPP_truepos_results |> dplyr::inner_join(deltaTm_truepos |> dplyr::select(Protein,delta_Tm),by="Protein")
```

```{r}
TPP_truepos_results |> dplyr::group_by(ATE_bin) |> dplyr::summarise(N=dplyr::n(),sens = mean(p_NPARC<0.05)) |> dplyr::ungroup() |>
  dplyr::mutate(spec = mean(TPP_trueneg_results$p_NPARC>0.05), prevalance = 0.10,
                PPV = (sens*prevalance)/((sens*prevalance) + (1-spec)*((1-prevalance))))
```



```{r}
TPP_truepos_results |> dplyr::mutate(delta_Tm_bin=cut(delta_Tm,breaks=c(-4,-1,0,1,2,3,5))) |> dplyr::group_by(delta_Tm_bin) |> dplyr::summarise(N=dplyr::n(),sens = mean(p_NPARC<0.05)) |> dplyr::ungroup() |>
  dplyr::mutate(spec = mean(TPP_trueneg_results$p_NPARC>0.05), prevalance = 0.10,
                PPV = (sens*prevalance)/((sens*prevalance) + (1-spec)*((1-prevalance))))
```

## Stratify by Corr

```{r}

MSstat_trueneg_results = sigmoid_trueneg_MSstat_result$ComparisonResult |> dplyr::inner_join(sigmoid_truenegative_simdata$params,by="Protein")


MSstat_truepos_results_bycorr  = MSstat_truepos_results |> dplyr::group_by(rho) |> dplyr::summarise(N=dplyr::n(),sens = mean(pvalue<0.05)) 

MSstat_trueneg_results_bycorr  = MSstat_trueneg_results |> dplyr::group_by(rho) |> dplyr::summarise(N=dplyr::n(),spec = mean(pvalue>0.05)) 

MSstat_results_bycorr = MSstat_truepos_results_bycorr |> dplyr::inner_join(MSstat_trueneg_results_bycorr,by="rho") |> dplyr::select(-N.y) |> dplyr::mutate(prevalance = 0.10,
                PPV = (sens*prevalance)/((sens*prevalance) + (1-spec)*((1-prevalance))))

MSstat_results_bycorr

```

```{r}

TPP_trueneg_results = TPP_trueneg_results |> dplyr::rename(Protein=uniqueID) |> dplyr::inner_join(sigmoid_truenegative_simdata$params,by="Protein")

```



```{r}

TPP_truepos_results_bycorr = TPP_truepos_results |> dplyr::group_by(rho) |> dplyr::summarise(N=dplyr::n(),sens = mean(p_NPARC<0.05)) 

TPP_trueneg_results_bycorr  = TPP_trueneg_results |> dplyr::group_by(rho) |> dplyr::summarise(N=dplyr::n(),spec = mean(p_NPARC>0.05)) 


TPP_results_bycorr = TPP_truepos_results_bycorr |> dplyr::inner_join(TPP_trueneg_results_bycorr,by="rho") |> dplyr::select(-N.y) |> dplyr::mutate(prevalance = 0.10,
                PPV = (sens*prevalance)/((sens*prevalance) + (1-spec)*((1-prevalance))))

TPP_results_bycorr

```

